HEP Software Foundation Community White Paper Working Group -- Training, Staffing and Careers

### 2018-07-04 CWP/papers/HSF-CWP-2017-02_training/latex/cwp-wg-training-careers.tex
Training last points (#97)
Commented by John and Graeme make the arguments more concise.

Reorganise the material in Sec 4 to present things in a more
coherent way, organised by topic. Reduce duplication of points
in places.

Remove "Roadmap" as a section 4 title as this is not
really what's prestented. Point to continuing work in the
auspices of the HSF Training WG in the conclusion.

Add some references and correct a few last points.

\begin{itemize}
    \item to encourage and provide incentives to the members of the HEP 
    community to train students and other collaborators,
    \item to properly assign credit to software development as a scientific 
    discipline, including training activities,
    \item to establish policies for the hiring and long-term retention of researchers 
    specialising computing,
    \item to address the gap in formal software training that is not always 
    given by universities as part of a physicist's education.
\end{itemize}

@@

A key goal is to find incentives that will encourage 
increasing numbers of people to dedicate their time and effort to train their
colleagues.

To be effective, the training must be done by people who work at the cutting
edge of the technology, who in many cases are found among the youngest
researchers. These are also the very same people who are most in need of
officially recognized credits for the advancement of their careers. Currently
visibility and recognition is given mainly to those working on data analysis
projects, rather than to the development and support of the underlying
software. One solution to this problem would be the establishment of specific
career paths for researchers specialising in scientific software development. 
%% GS Comment - is this actually ``motivation'' - it seems more like a solution
% See suggested title change for this section...
However, the practical 
implementation of such an approach is extremely challenging, as there are a wide range of
institutions involved that belong to different countries with their own
policies, priorities, and funding strategies.

Broadly speaking, two scientific profiles of researchers could be
envisaged; a physicst could make a detailed plan of what they need, expressed 
in a tidy requirements document, and a computer scientist could use this document 
to provide the required solution. This is how the software development process 
typically works in the software industry. However, in HEP requirements are rapidly
changing, forcing developers and physicsts to interact closely during the entire
process of software development; this works smoothly only when both communities have the
same goals and speak the same language, hence the need for physicists with a good
knowledge of computer science.

The first possible career would be the path of a \emph{physicist researcher with
computing science specialisation}, in some places corresponding to the role of a
physicist-programmer. It would be conceptually different to that of a
\emph{computer scientist}: the two have different goals, seek different paths to
the solution of their problems, and usually do not even adopt the same language in
expressing problems and solutions. Such a role would have an active role in
physics analysis, and so meet many of the traditional metrics for an academic
career path.

The second would be the path of a domain-specific \emph{researcher software
engineer}. Such a person has the primary role of developing software and finding
software solutions, but with a large domain-specific knowledge to both
understand the use cases and clients, but also of the available solutions.  Such
posts are almost impossible to develop in the current academic frameworks; but
while RSEs are understood by many funding agencies and universities as a
technical post, they are seen as available for short-term engagements with a
researcher community, delivering a quick fix or a refactoring and then moving
on. This does not deliver the domain specific expertise required.

Both career paths suffer in academic terms from few or poorly cited
publications, few opportunities to win grants in their own right and little
opportunity for impact with commerce and wider society. These three failings
make an academic appointment challenging or disfavoured compared to other
applicants. However, there are measures that can be taken to address all three
weaknesses.
%% GS Comment - This is an odd place to end a section. What are these measures?
% Where do we list them? Even a forward reference would help.

% This makes the development of specific solutions even more difficult and
% cumbersome.

%Finally, the career of these two figures follows slightly different paths, with different requirements. 
%____________________________________________________________________
%\subsection{Existing Training efforts  \textcolor{pink}{is this needed?....}}
% why not, shows what is done already

@@

% DOES THIS SECTION ANSWER THE FOLLOWING: 
% - Different levels of training are required

The HEP user community consists of a set of users with diverse software
experience, interests, and time availablity to learn new techniques.  Any
training programme must take into account the variation in target audiences. For
example, an undergraduate doing a summer research project has different needs
and skills than their professor.

@@

For purposes of training, it is useful to think of all our user community as
\emph{students}. More specifically, we can broadly identify four different sets
of students, classified by experience level:

\begin{itemize}
   \item \emph{Beginners}: New collaborators with no knowledge of the 
   tools or techniques they are expected to use. These might be undergraduates, 
   but they could also be senior researchers who have not yet, or not recently, 
   engaged in software analysis. These are people in need of some kind of formal 
   training in modern computing techniques (basically compiled and scripting 
   languages, together with operating system basics such as filesystems, version 
   control, and command-line shells).
   \item \emph{Intermediate level analysts}: people with some experience in 
   either analysis concepts and/or tools, but  looking to supplement their 
   experience with more recent and modern approaches. 
   \item \emph{Advanced level analysts}: experts who have mastered current 
   technologies and implementations and who want to stay up-to-date with 
   new advanced developments.
   \item \emph{Software developers}: HEP scientists in charge of software 
   development in areas not limited only to analysis, such as DAQ systems, 
   computing infrastructure, databases, pattern recongition, and complete 
   frameworks.
\end{itemize}

@@

At all stages of software and computing training, we should take care to
encourage \emph{Good Practices Across the Community}, such as error
checking, modularity of code design, version control, writing tests, etc. All
the key concepts addressed in the training should not be specific to a
particular experiment or field of application, but general enough to be useful
to the whole HEP community and beyond. In this section, we present a
list of specific concepts that need to be taught, in
order to guarantee the basic level of competence needed to write efficient code
for the different tasks performed in HEP experiments.

Essential knowledge that needs to be transferred includes basic
programming concepts, data structures, basics of code design, error checking,
code management tools, validation and debugging tools. More advanced topics
include modularity of code design, advanced data structures, evaluation metrics,
writing tests and working with different types of hardware accelerator.
Additionally, special emphasis should be given to
reporting results and documenting them.

@@

% ____________________________________________________________________
% ____________________________________________________________________
\section{Implementation Roadmap}
The implementation of training should target multiple training formats and
knowledge transfer, such as videos, wikis, lectures, jupyter notebooks and
advanced visualizations, etc.
People who learn in non-conventional ways should not be excluded from the
opportunity to learn. 
%% GS Comment - I'm not really sure what that previous sentence is supposed to
% mean.
Since there are different stages of knowledge transfer,
programmes should be developed that encourage experts to share knowledge. Issues
considered in this section include global training initiatives, strategies for
leveraging existing forums, and training tool development.

An important point to consider is the difficulty, often experienced in the past,
in developing large software programs across different experiment communities.
While there already are experiment-specific training efforts in place, there are
many needs that are in common between experiments. Therefore, we should probably
strive to extract the common knowledge and build a common training programme. This
would reduce duplication of efforts and enable growth of a shared training
culture by accumulating and sharing expertise across different experiments.
To realise this goal, a possibility could be the institution of a {\em
federation} of existing training initiatives, a very lightwheight umbrella just
aimed at improving the efficiency and cost-effectivness of this important
activity. An appropriate incentive programme to reward those who train the community 
might help to facilitate the goals of such an initiative.

When trying to exchange training materials within a group the first problem is
to convince authors to contribute their work. This issue
can be alleviated with cautious intellectual property clauses. Another
problem is that trainers do not like to reuse given material as-is. They usually
want to refactor it, building their own training story.
This makes it difficult to have everyone agree on common material.
It is even challenging to agree to host material in the same
centralized place. A way to overcome this could be to settle for a centralized
catalog, with each author being free to host their material in their location
of choice.

Another challenge is the wide heterogeneity in
student competence. Special care must be given to setup
a training structure that can cope with both beginners and advanced students.
This may rest on a low granularity subdivision of independant topics.
The IN2P3 authors try to restrict their tutorials to 25 minutes
of material, as inspired by the Pomodoro technique\cite{PomodoroTechnique}, so
that one can easily jump and compose one's own ``menu''. However, it turns out to be tricky to keep such
small tutorials really independant and meaningful when they can be selected at
will.

All trainers have also faced the tremendous loss of time
in software installation for any practical exercises as many students have
not managed to prepare their machines in advance.
Here containers may bring real
progress, provided that everyone has at least Docker\cite{Docker} installed on
their machine.
JupyterHub\cite{JupyterHub}, also, is a technology which enable training session
with no need for installation (but with a need for network).

%____________________________________________________________________
\subsection{Initiatives for future training programmes\label{sec:initiatives}}

Some methods that can be used for global training include Massive Online Open
Courses (MOOCs), hands-on workshops, online knowledge bases, expert trainer volunteer
networks, and web-based training approaches.

MOOCs can be used to develop an open-source set of
tutorials and tools. Existing online courses such as Udacity~\cite{Udacity} and 
Coursera~\cite{Coursera} can be
evaluated and recommended to the community. A novel approach such as WikiToLearn~\cite{WikiToLearn}
could also be explored to assess potential benefits (as has already been attempted
in the context of the GridKa School of Computing~\cite{GridKa}, see below).

@@

Question-and-answer websites such as Stack Overflow~\cite{Stackoverflow} for HEP are also a very
useful resource for common problems and questions. This approach has already
been considered by HSF start-up members, and turned out to be difficult to
pursue for a number of reasons, but in the future, boundary conditions might
change, making this approach viable.

@@

Creating an expert tutor volunteer network is another way to provide training
and support to the community.
This of course requires proper recognition, at least in terms of career
prospects as an incentive for young researchers. It is clear that the best
possible tutors are, in principle, those people who are actively engaged in
modern software developments: these are often young researchers in the
first steps of their careers and, in order to be attracted to devoting
substantial time to training and tutoring, they must be assured that such an
effort would be properly recognised in an official way. Such kind of recognition
is not currently implemented, at least not in a standardized or official way. A
possible structure for such a network could be the establishment of a 
{\em federation} of existing schools.

It should be considered that some professors, who act as graduate student
advisors, might need to be encouraged to make sure their students are properly
trained. Sometimes, students are instead pushed to learn the bare minimum to get
the work done, at the expense of a broader training/education curriculum that
could actually yield improved results further down the line. One incentive
would be to provide training programmes that also count as course credit, perhaps
as an elective. This model is already in limited use with some online
solutions~\cite{TrainingProgram}, but this is not universal.
Discussions should be started with collaborators at higher education
institutions to see what the roadblocks or opportunities would be for these
training sessions to serve double duty.
It should also be considered that not all students will end their career in
research or academia: their contribution to the research activity, as students,
should therefore also provide them knowledge, know-how, and skills considered a
valuable asset by industry, in order to increase their chances of a career
outside research.

Difficulties that have emerged in the past with respect to implementing training
courses are the lack of funding and the lack of available time by experts in the
field. People with enough expertise or insight usually don't have
time to devote to prolonged periods of student training, and, even when they can
find time, the cost of setting up a training course in an effective way is often
beyond what's made available by funding agencies (funds for travel, hosting,
setting up a room with a computing infrastructure to allow interactive hands-on
session, etc.) A possible solution is a completely different approach to
training (but complementary to the already existing and successful classical
efforts such as the CERN School of Computing's Bertinoro and KIT ones): instead
of directly teaching to students, trainees could make use of a web-based
platform to provide training materials to students. This complementary approach
has several advantages over traditional ones:
\begin{itemize}
   \item Tutors can add material to the web site at a very slow pace (whenever
   they find time to do it, one slide a week or a chapter a day) 
   \item Their material, available public on a web-site, can be further expanded
   by collaborators (also at their own pace) or, even better and more productively, by students
   who decide to contribute new additions, examples, exercises etc. Such a
   collaborative effort allows more people to be exposed to training at any
   given time, creates a sense of community, and creates bridges between people
   in contiguous areas of research. Students can use the same platform to
   exchange their own examples, make suggestions and point out interesting
   concepts. In such a model, the possible contribution from others to the
   training material needs to be moderated and validated with appropriate
   policies.
   \item If complemented by the availability of remote virtual machines
   (possibly through a browser), students could have access to examples and
   exercises that are already embedded in their own natural environment: all the
   necessary tools/libraries needed to implement the exercise will be already
   available in the virtual machine (possibly a Docker container). With just a
   web browser, students could run complex examples from home, taking advantage
   of a remote facility that provides some storage and computing power.
   Important here is the concept of ``environment'': a Docker container could be
   set up in such a way that students will work in an exact replica of the
   environment they will be exposed to in their experiment. Moreover, students 
   could be provided with Docker containers that
   preserve their modified environment across sessions, allowing them to
   develop their skills over a prolonged period of time by accessing all the
   files that were made persistent day by day during the training.
   \item There would no longer be a need to find the resources to host a school
   and pay the tutor(s) (and eventually subsidize the students to participate in
   training in a remote location). Students could follow the training at their
   own pace from wherever they happen to be. A traditional school only lasts for
   5 days (10 at most) and it is difficult to cover a  subject to any
   significant depth in such a short time. The web approach, instead, would
   allow for very long and in-depth coverage of any kind of subject, and in this
   sense it could be a {\it complementary} approach to a traditional school. Of
   particular interest could be courses such as ``Machine Learning'',
   ``Statistical Analyis with ROOT'', or even just ``Good practices in C++'' or
   ``Python Programming for scientific computing''.
   \item Finally, this approach could allow the creation of {\it browsable}
   repositories of all training materials, grouped by argument, by relevance, by
   experiment or whatever other critieria. Students from all over the world
   could be exposed to a large repository of examples, exercises and in general
   training material from their own home.
\end{itemize}

@@

The adoption of such an approach is made rather easy in WikiToLearn by the
relative simplicity of the wikimedia-based toolset: users contribute their
training material using just a web-browser, and in order to do this efficiently,
the necessary learning curve has been kept appropriately shallow.
An interesting exercise in this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL~\cite{WikiToLearnGridka}.
This constitutes an interesting example of what can be accomplished using this
platform; it is just a first example of what's possible, but an inspiring one.

@@

  % \item Question-and-answer web sites such as Stack Overflow 
   %(https://en.wikipedia.org/wiki/Stack_Overflow) 
   %/ Stack Exchange are very useful. 
   %(https://en.wikipedia.org/wiki/Stack_Exchange)
  % \item An expert/tutor volunteer network (office hours style) can be established.
   %\item Investing in periodic in-person, hands-on workshops can be continued.

%____________________________________________________________________
\subsection{Leveraging existing forums}

@@

% Anecdotally, these are viewed as a great success by the CMS community in
% preparing new collaborators and might serve as a model for future efforts in
% ML training.

@@

There are a growing set of MOOCs teaching various subjects. Since there are many
options, there is a wide variety with respect to the depth of the material and
specific tools taught. Exploring these options allows us to choose which is the
right offering for the knowledge needed to work on a specific experiment. We
can pick and choose modules to tailor an appropriate roadmap of skills to learn.
More difficult will be to assemble specific training material not already
avaible elsewhere in an efficient and organized way, since this requires
adequate organization, volounteers and a suitable infrastructure.

@@

@@

The HEP community by and large acknowledges and recognizes the great importance
of training in the field of scientific computing. This activity should encompass
several types of {\it students}, from undergraduates, to young researchers, up
to senior physicists, all of them in need of an appropriately designed training
path in order to be proficent in their scientific endeavours.

@@

For each of these points we have provided an overview of the current situation
and made proposals to improve the situation in HEP.

%% GS Comment - this section is incomplete. FIXME.

>>>>

\begin{itemize}
    \item to encourage and provide incentives to the members of the HEP 
    community to train students and other collaborators,
    \item to properly assign credit to software development as a scientific 
    discipline, including training activities,
    \item to establish policies for the hiring and long-term retention of researchers 
    specialising in computing,
    \item to address the gap in formal software training that is not always 
    given by universities as part of a physicist's education.
\end{itemize}

@@

A key goal is to find incentives that will encourage 
increasing numbers of people to dedicate their time and effort to train their
colleagues.
To be effective, the training must be done by people who work at the cutting
edge of the technology and who in many cases are found among the youngest
researchers. These are also the very same people who are most in need of
officially recognized credits for the advancement of their careers. Currently,
visibility and recognition is given mainly to those working on data analysis
projects, rather than to the development and support of the underlying
software. One solution to this problem would be the establishment of specific
career paths for researchers specialising in scientific software development. 
However, the practical 
implementation of such an approach is extremely challenging, as there are a wide range of
institutions involved that belong to different countries with their own
policies, priorities, and funding strategies.

Broadly speaking, two scientific profiles of researchers could be
envisaged. A physicst could make a detailed plan of what they need, expressed 
in a tidy requirements document, and a computer scientist could use this document 
to provide the required solution. This is how the software development process 
typically works in the software industry. However, in HEP requirements are rapidly
changing, forcing developers and physicsts to interact closely during the entire
process of software development. This works smoothly only when both communities
have the same goals and speak the same language, hence the need for physicists with a good
knowledge of computer science.

The first possible career path would be that of a \emph{physicist with computing
science specialisation}, also known as a physicist-programmer in some
communities. (It would be conceptually different to that of a \emph{computer
scientist} as the two have different goals, seek different paths to the solution
of their problems, and usually do not even share the same language.) Such people
would have an active role in physics analysis, and so meet many of the criteria
needed for an academic career path.

The second would be the path of a domain-specific \emph{software
engineer}. Such a person has the primary role of developing software and finding
software solutions, but with a large domain-specific knowledge to
understand relevant use cases and the available solutions. 
Such posts are almost impossible to establish in current academic
institutions; rather being seen by many funding agencies and
universities as short term technical positions. This does not allow domain
specific expertise to be aquired or retained in our field.

Both career paths suffer in academic terms from few or poorly cited
publications, few opportunities to win grants in their own right and little
opportunity for impact with commerce and wider society. These three failings
make an academic appointment challenging or disfavoured as compared to other
applicants. However, there are measures that can be taken to address all three
weaknesses, which we expand on in the remainder of this document.

@@

The HEP community consists of people with diverse software
experience, interests, and time availablity to learn new techniques.  Any
training programme must take into account the variation in target audiences. For
example, an undergraduate doing a summer research project has different needs
and skills than their professor.

@@

For purposes of training we can broadly identify classify four
different experience levels:

\begin{itemize}
   \item \emph{Beginner}: New collaborators with no knowledge of the 
   tools or techniques they are expected to use. These are people in need of some kind of formal 
   training in modern computing techniques, such as compiled and scripting 
   languages, together with operating system basics such as filesystems, version 
   control, and command-line shells.
   \item \emph{Intermediate}: People with some experience in 
   concepts and tools, but looking to supplement their 
   experience with more recent and modern approaches. 
   \item \emph{Advanced}: Experts who have mastered current 
   technologies and implementations and who want to stay up-to-date with 
   new advanced developments.
   \item \emph{Software Specialists}: HEP scientists in charge of software 
   development in areas not limited only to analysis, such as DAQ systems, 
   computing infrastructure, databases, pattern recongition, and complete 
   frameworks.
\end{itemize}

@@

At all stages of software and computing training, we should take care to
encourage \emph{good practices across the community}, such as ensuring error
checking, modularity of code design, version control, writing tests, etc. All
the key concepts addressed in training should not be specific to a particular
experiment or field of application, but general enough to be useful to the whole
HEP community and beyond. A number of specific concepts need to be taught, in
order to guarantee the basic level of competence needed to write efficient code
for the various tasks that need to be performed in HEP experiments.
These include
programming concepts, data structures, basics of code design, error checking,
code management tools, validation and debugging tools. More advanced topics
include modularity of code design, advanced data structures, evaluation metrics,
writing tests and working with different types of hardware accelerator.
Additionally, special emphasis should be given to
reporting results and documenting them.

@@

% ____________________________________________________________________
% ____________________________________________________________________
\section{Advancing Training in HEP}

The implementation of training should employ different training formats, such as
videos, wikis, lectures, jupyter notebooks and advanced visualizations, etc.,
so that people can learn in a familiar and effective manner and in such a way
that experts are encouraged to share their knowledge.

An important point to consider is the difficulty, often experienced in the past,
in developing large software programs across different experiment
collaborations.
While there already are experiment-specific training efforts in place, there are
many needs that are in common. Establishing a common training programme could
help facilitate the sharing of exprience amongst different experiments. This
would reduce duplication of efforts and enable growth of a shared training
culture by accumulating and sharing expertise.
To realise this goal, a possibility could be the creation of a 
\emph{federation} of training initiatives,
aimed at improving the efficiency and cost-effectivness of this important
activity. An appropriate incentive programme to reward those who train the community 
might help to facilitate the goals of such an initiative.

When trying to exchange training materials within a group the first problem is
to convince authors to contribute their work. This issue
can be alleviated by addressing the correct assignment of intellectual property.
Another problem is that trainers do not like to reuse given material as-is. They usually
want to refactor it, building their own training history.
This makes it difficult to have everyone agree on common approaches.
It is even challenging to agree to host material in the same
centralized place. A way to overcome this could be to settle for a centralized
catalog, with each author being free to host their material in their location
of choice.

Another challenge is the wide range of student competence. Special care must be
given to setup a training structure that can manage both introductory and
advanced material. This may be addressed by organising course material as a
large number of independant topics. The IN2P3 authors try to restrict their
tutorials to 25 minutes, as inspired by the Pomodoro
technique\cite{PomodoroTechnique}, so that one can easily jump and compose one's
own curriculum. However, it turns out to be tricky to keep such small tutorials
really independant and meaningful when they can be selected at will.

All trainers have also faced the tremendous loss of time
in software installation for any practical exercises as many students have
not managed to prepare their machines in advance.
Here containers may bring real
progress, provided that everyone has at least Docker\cite{Docker} installed on
their machine.
JupyterHub\cite{JupyterHub} is a technology which supports training sessions
without the need for specilaist installations.

%____________________________________________________________________
\subsection{Initiatives for Future Training Programmes\label{sec:initiatives}}

Some methods that can be used for location independent training include Massive
Online Open Courses (MOOCs), hands-on workshops, online knowledge bases, expert trainer volunteer
networks, and web-based training approaches.

MOOCs can be used to develop an open-source set of
tutorials and tools. Existing online courses such as Udacity~\cite{Udacity} and 
Coursera~\cite{Coursera} can be
evaluated and exploited by the community. A novel approach such as
WikiToLearn~\cite{WikiToLearn} could also be explored to assess potential benefits (as has already been attempted
in the context of the GridKa School of Computing~\cite{GridKa}, see below).

@@

Question-and-answer websites such as Stack Overflow~\cite{Stackoverflow} for HEP are also a very
useful resource for common problems and questions. This approach has already
been considered by HSF start-up members, and turned out to be difficult to
pursue due to the lack of a critical mass needed by Stack Overflow, but in the
future, boundary conditions might change, making this approach viable.

@@

Creating an expert tutor volunteer network is another way to provide training
and support to the community.
This of course requires proper recognition, at least in terms of career
prospects as an incentive for young researchers. It is clear that the best
possible tutors are, in principle, those people who are actively engaged in
modern software developments: these are often young researchers in the
first steps of their careers and, in order to be attracted to devoting
substantial time to training and tutoring, they must be assured that such an
effort would be properly recognised in an official way. Such kind of recognition
is not currently implemented, at least not in a standardized or official way. A
possible structure for such a network could be the establishment of a 
{\em federation} of existing schools, as discussed in \S\ref{sec:ehcurtrainprog}.

\subsection{Web Based Training}

Difficulties that have emerged in the past with respect to implementing training
courses are the lack of funding and the lack of available time by experts in the
field. People with enough expertise or insight usually don't have time to devote
to prolonged periods of student training, and, even when they can find time, the
cost of setting up a training course in an effective way is often beyond what is
made available by funding agencies (funds for travel, hosting, setting up a room
with a computing infrastructure to allow interactive hands-on session, etc.) A
possible solution is a completely different approach to training, using a
web-based platform to provide training materials to students. This would be
complementary to the already existing and successful efforts such as the CERN
School of Computing's Bertinoro and KIT ones.

The web based approach has several advantages over traditional ones:
\begin{itemize}
   \item Tutors can add material to the web site at a very slow pace (whenever
   they find time to do it, one slide a week or a chapter a day) .
   \item Their material, available public on a web-site, can be further expanded
   by collaborators (also at their own pace) or, even better and more productively, by students
   who decide to contribute new additions, examples, exercises etc. Such a
   collaborative effort allows more people to be exposed to training at any
   given time, creates a sense of community, and creates bridges between people
   in contiguous areas of research. Students can use the same platform to
   exchange their own examples, make suggestions and point out interesting
   concepts. In such a model, the possible contribution from others to the
   training material needs to be moderated and validated with appropriate
   policies.
   \item If complemented by the availability of remote virtual machines
   (possibly via a browser), students could have access to examples and
   exercises that are already embedded in their own natural environment: all the
   necessary tools and libraries needed to implement the exercise will be
   already available in the virtual machine (possibly a Docker container). With just a
   web browser, students could run complex examples from home, taking advantage
   of a remote facility that provides some storage and computing power.
   Important here is the concept of ``environment'': a Docker container could be
   set up in such a way that students will work in an exact replica of the
   environment they will be exposed to in their experiment. Moreover, students 
   could be provided with Docker containers that
   preserve their modified environment across sessions, allowing them to
   develop their skills over a prolonged period of time by accessing all the
   files that were made persistent day by day during the training.
   \item There would no longer be a need to find the resources to host a school
   and pay the tutor(s) (and eventually subsidize the students to participate in
   training in a remote location). Students could follow the training at their
   own pace from wherever they happen to be. A traditional school only lasts for
   5 days (10 at most) and it is difficult to cover a  subject to any
   significant depth in such a short time. The web approach, instead, would
   allow for very long and in-depth coverage of any kind of subject, and in this
   sense it could be a {\it complementary} approach to a traditional school. Of
   particular interest could be courses such as ``Machine Learning'',
   ``Statistical Analyis with ROOT'', or even just ``Good practices in C++'' or
   ``Python Programming for scientific computing''.
   \item Finally, this approach could allow the creation of {\it browsable}
   repositories of all training materials, grouped by argument, by relevance, by
   experiment or whatever other critieria. Students from all over the world
   could be exposed to a large repository of examples, exercises and in general
   training material from their own home.
\end{itemize}

@@

The adoption of such an approach is made rather easy in WikiToLearn by the
relative simplicity of the wikimedia-based toolset: users contribute their
training material using just a web-browser, and in order to do this efficiently,
the necessary learning curve has been kept appropriately shallow.
An interesting exercise in this context has recently been made by colleagues of
the GridKa School of computing: the material from this year (2017) has been made
publicly available on WTL~\cite{WikiToLearnGridka}.
This constitutes an interesting example of what can be accomplished using this
platform; it is just a first example of what is possible, but an inspiring one.

@@

%____________________________________________________________________
\subsection{Enhancing Current Training Programmes}
\label{sec:ehcurtrainprog}

@@

@@

There are a growing set of MOOCs teaching various subjects. Since there are many
options, there is a wide variety with respect to the depth of the material and
specific tools taught. Exploring these options allows us to choose which is the
right offering for the knowledge needed to work on a specific experiment. We
can pick and choose modules to tailor an appropriate roadmap of skills to learn.
More difficult will be to assemble specific training material not already
avaible elsewhere in an efficient and organized way, since this requires
adequate organization, volunteers and a suitable infrastructure.

@@

It should be considered that some graduate student
advisors, might need to be encouraged to make sure their students are properly
trained. Sometimes, students are instead pushed to learn the bare minimum to get
the work done, at the expense of a broader training/education curriculum that
could actually yield improved results further down the line. One incentive
would be to provide training programmes that also count as course credit, perhaps
as an elective. This model is already in limited use with some online
solutions~\cite{TrainingProgram}, but this is not universal.
Discussions should be started with collaborators at higher education
institutions to see what the roadblocks or opportunities would be for these
training sessions to serve double duty.
It should also be considered that not all students will end their career in
research or academia: their contribution to the research activity, as students,
should therefore also provide them knowledge, know-how, and skills considered a
valuable asset by industry, in order to increase their chances of a career
outside research.

@@

The HEP community by and large acknowledges and recognizes the great importance
of training in the field of scientific computing. This activity should encompass
several types of \emph{students}, from undergraduates, to young researchers,
up to senior physicists, all of them in need of an appropriately designed training
path in order to be proficent in their scientific endeavours.

@@

For each of these points we have provided an overview of the current situation
and made proposals to improve the situation in HEP. The ideas presented need to
be developed further and concrete actions in the community need to be
implemented, which we will undertake in the HSF Training Working Group\cite{HSFTrainingWG}.
